# Special formatting {#format}

In development.

## Formatting

### Columns

As in the presentations, we are also able to distribute content accross columns here. While we would do this in the presentations using `.pull-right[]` or `.pull-left[]`, here we will have to use the HTML content division element, which is opened by `<div>` and closed by `</div>`.

There are current several different classes of content division that we implemented in this book and that we can specify by adding `class` to the content division element opener. For instance, to add the `two-col` class to `<div>`, we would do the following: `<div class = 'two-col'>`.

Remember that we must always close the HTML `<div>` by including `</div>` at the end of the content we wanted to divide; there is no need to include `class` when closing any HTML elements.

We list below all examples of content dividers that can be used in this book.

#### The simplest two-column split: class `two-col`

To equally divide the content across two equal-sized columns, we include it between `<div class = 'two-col'>` `</div>`. 

<div class = 'two-col'>

1. Introduction à R
2. Charger et manipuler des données
3. Introduction à ggplot2
4. Modèles linéaires
5. Programmation avec R
6. Modèles linéaires généralisés
7. Modèles generales et generalisées linéaires à effets mixtes
8. Modèles additifs généralisés
9. Analyses multivariées
10. Analyses multivariées avancées

</div>

As shown below:

```
<div class = 'two-col'>
1. Introduction à R
2. Charger et manipuler des données
3. Introduction à ggplot2
4. Modèles linéaires
5. Programmation avec R
6. Modèles linéaires généralisés
7. Modèles generales et generalisées linéaires à effets mixtes
8. Modèles additifs généralisés
9. Analyses multivariées
10. Analyses multivariées avancées
</div>
```

#### The simplest three-column split: class `three-col`

To equally divide the content across two equal-sized columns, we include it between `<div class = 'three-col'>` `</div>`. 

<div class = 'three-col'>

1. Introduction à R
2. Charger et manipuler des données
3. Introduction à ggplot2
4. Modèles linéaires
5. Programmation avec R
6. Modèles linéaires généralisés
7. Modèles generales et generalisées linéaires à effets mixtes
8. Modèles additifs généralisés
9. Analyses multivariées
10. Analyses multivariées avancées

</div>

As shown below:

```
<div class = 'three-col'>
1. Introduction à R
2. Charger et manipuler des données
3. Introduction à ggplot2
4. Modèles linéaires
5. Programmation avec R
6. Modèles linéaires généralisés
7. Modèles generales et generalisées linéaires à effets mixtes
8. Modèles additifs généralisés
9. Analyses multivariées
10. Analyses multivariées avancées
</div>
```

> Unfortunately, the `two-col` and `three-col` classes are not ideal to split code and output across two distinct columns. See other options below.

#### More flexible split across two columns

##### Even two-column split: classes `split`, `split-left`, and `split-right`

A more flexible way of splitting content across two columns is to use the `split`, `split-left`, and `split-right` classes. `split` will create a `flex` box layout-container and `split-left` and `split-right` can be used to add content to this layout. See an example of its usage below:

> This is the recommended option if you would like to split code and prose across two evenly-sized columns. 

<div class = "split">
<div class = "split-left">
**Introduction to standard deviation**

Let us calculate the standard deviation for a population containing the values $1$, $2$, $3$, $4$, and $5$.

We can calculate the standard deviation using the `sd()` function in `R`, as shown below:
```{r sd-example, echo = TRUE, results = "hide"}
sd(c(1, 2, 3, 4, 5))
```

</div>

<div class = "split-right">
Remember that the **standard deviation** can be calculated as: 

$$\sigma = \sqrt{\frac{\sum\limits_{i=1}^{n} \left(x_{i} - \bar{x}\right)^{2}} {n-1}}$$

The output for the code on the left side is:

```{r sd-example-output, echo = FALSE, ref.label='sd-example'}
```

</div>
</div>

The above example has been generated using:

````markdown
<div class = "split">
<div class = "split-left">
**Introduction to standard deviation**

Let us calculate the standard deviation for a population containing the values $1$, $2$, $3$, $4$, and $5$.

In `R`, we could do the following:
`r ''````{r sd-example, echo = TRUE, results = "hide"}
    sd(c(1, 2, 3, 4, 5))
```

</div>
<div class = "split-right">

Remember that the **standard deviation** can be calculated as: 
$$\sigma = \sqrt{\frac{\sum\limits_{i=1}^{n} \left(x_{i} - \bar{x}\right)^{2}} {n-1}}$$

The output for the code on the left side is:
    
`r ''````{r sd-example-output, echo = FALSE, ref.label='sd-example'}
```
</div>
</div>
````

##### Uneven two-column uneven-split: classes `split`, `uneven-split-left`, and `uneven-split-right`

To use uneven columns, use the `uneven-split-left`, and `uneven-split-right` classes inside the `split` class.See an example of its usage below:

<div class = "split">
<div class = "uneven-split-left">
**Introduction to standard deviation**

Let us calculate the standard deviation for a population containing the values $1$, $2$, $3$, $4$, and $5$.

We can calculate the standard deviation using the `sd()` function in `R`, as shown below:
```{r sd-example-2, echo = TRUE, results = "hide"}
sd(c(1, 2, 3, 4, 5))
```

</div>

<div class = "uneven-split-right">
Remember that the **standard deviation** can be calculated as: 

$$\sigma = \sqrt{\frac{\sum\limits_{i=1}^{n} \left(x_{i} - \bar{x}\right)^{2}} {n-1}}$$

The output for the code on the left side is:

```{r sd-example-output-2, echo = FALSE, ref.label='sd-example'}
```

</div>
</div>

The above example has been generated using:

````markdown
<div class = "split">
<div class = "uneven-split-left">
**Introduction to standard deviation**

Let us calculate the standard deviation for a population containing the values $1$, $2$, $3$, $4$, and $5$.

In `R`, we could do the following:
`r ''````{r sd-example-2, echo = TRUE, results = "hide"}
    sd(c(1, 2, 3, 4, 5))
```

</div>
<div class = "uneven-split-right">

Remember that the **standard deviation** can be calculated as: 
$$\sigma = \sqrt{\frac{\sum\limits_{i=1}^{n} \left(x_{i} - \bar{x}\right)^{2}} {n-1}}$$

The output for the code on the left side is:
    
`r ''````{r sd-example-output-2, echo = FALSE, ref.label='sd-example'}
```
</div>
</div>
````

### Special notes and tips

Standard Markdown notes can be obtained by typing `>` at the beginning of the paragraph followed by space.

> For example, like this.

Here, we can also use special note tips by writing `:::` followed by one of the four class names available: `puzzle`, `explanation`, and `noway`.

To make special explanation boxes, use `explanation`:

:::explanation

You use a div tip by writing `:::` following by the name that you assigned to it in the CSS after the `div`.

:::

To pose thoughtful questions to the readers, use `puzzle`:

:::puzzle
What is the spell that allows wizards to make objects levitate with the flick of their wand?
:::

To tell readers they must not do something, use `noway`:

:::noway
You may not begin function names in `R` with numbers.
:::


To show that you are not happy about something, use `angry`:

:::angry
Do not really use this. This is just a test.
:::
