# References and citations

In development.

We already covered the basics on how to insert an inline citation on Chapter \@ref(basics). However, those examples only work for this book because they are the packages we are using to create the book. If we wanted to cite another package or a book, article, we first have to create the bibliography entries for them.

## The BibTeX format
Bookdown uses `pandoc-citeproc` for its html output. Many bibliography formats can be used by `pandoc-citeproc`, but [BibTex](http://www.bibtex.org/) is quite simple and relatively standard so we will use it for our books.

Here is a basic example of a BibTex entry:
```{r eval=FALSE}
@Book{abramowitz+stegun,
 author    = "Milton {Abramowitz} and Irene A. {Stegun}",
 title     = "Handbook of Mathematical Functions with
              Formulas, Graphs, and Mathematical Tables",
 publisher = "Dover",
 year      =  1964,
 address   = "New York City",
 edition   = "ninth Dover printing, tenth GPO printing"
}
```

Now let's see what the structure of the entry is:

- Each entry starts with a type (e.g. `@Book{`) such as `article` or `book`, among others.
- It is then followed by a citation-key (e.g. `abramowitz+stegun,`) which can be used to produce the inline citations. So if we wanted to cite our example book we would use `@abramowitz+stegun` to produce this citation: @abramowitz+stegun.
- Finally, the entry has multiple tags which give the characteristics of the entry (e.g. author, title, year, etc.)
  - There are some required tags for each type of entry as well as some optional ones.

You can find a list of the different types and the required and optional tags for each type [here](https://en.wikipedia.org/wiki/BibTeX).

## Creating a BibTeX database

Now that we know the basics about BibTeX, we can apply it. For this, we have to create a BibTeX database which is a plain-text file with a `.bib` extension. For example, we created an `example.bib`file so that we could use the @abramowitz+stegun citation.

The file name for the database of each workshop should be `workshop-X-citations.bib`.

### R packages

If you wish to cite any R packages, **knitr** contains the helper function `write_bib()` that automatically produces the BibTeX entries for the selected packages and creates a datbase for the packages. Here is an example that this template uses to produce citations such as this one for **bookdown** [@R-bookdown]:
```{r eval=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

## Adding the BibTeX database

To include the database in the book, you must:

- Save the `.bib` file in the `book-en` or `book-fr` folder.
- Add the file name to the `bibliography` field in the YAML metadata section of the `index.Rmd` file.
  - Multiple database files can be included. For example, the `bibliography` field of this template looks like this:
```{r eval=FALSE}
bibliography: [packages.bib, example.bib]
```

## The `XX-references.Rmd` file

The basic format of this Rmd file should only contain the following expression:
```{r eval=FALSE}
`r if (knitr::is_html_output()) '# Additional resources {-}'`

### References {-}
```

This will output only the references we included in the main BibTeX database (i.e. the `workshop-X-citations.bib` file) and any packages we cited inside the book.

- Note that this will not include the rest of the package reference entries created by **knitr** if they were not cited inside the book.

### Including other resources

Most of the wikis for the QCBS R Workshop Series contain a section at the end where we offer additional resources for the participants interested in more information about the subject of the workshop.

To include these into this last chapter of the book, you can add a new subsection after the basic expression showed above. In this book, we have included a link to the **bookdown** manual as an additional resource like this:
```{r eval=FALSE}
`r if (knitr::is_html_output()) '
# Additional resources {-}
'`

For more information on how to use bookdown, you can check the bookdown manual [here](https://bookdown.org/yihui/bookdown/).

### References {-}
```
